### 2024-12-02 paper/paper.qmd
Updated paper to include data section

---
title: "Analyzing the Pakistan Cricket Team's Performance (2002-2024)"
subtitle: "My subtitle if needed"
author: 
  - Muhammad Abdullah Motasim
thanks: "Code and data are available at: [https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data](https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data)."
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
table-of-contents: true
toc-depth: 2
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(ggplot2)
library(arrow)
library(patchwork)

pakistan_match_data <- read_parquet(here("data/02-analysis_data/cleaned_pakistan_match_data.parquet"))
```

# Introduction
Pakistan’s cricket team holds a storied legacy, marked by remarkable highs and challenging lows. From their 1992 World Cup triumph to fluctuations in form over the decades, analyzing trends in performance offers insights into their journey and areas for strategic improvement.

Despite an extensive focus on player statistics and match highlights, there remains a lack of systematic statistical analysis to evaluate long-term performance trends. This paper fills this gap by analyzing data from 2002 to 2024 to understand what factors contribute to Pakistan’s success in international cricket.

We employ a linear regression model to analyze match outcomes, considering variables like year, opponent, and toss outcomes. The focus is on identifying trends and predicting the probability of victory under varying conditions. Our findings not only contribute to the literature on sports analytics but also provide actionable insights for coaches, analysts, and cricket administrators not only in Pakistan, but around the world.

The structure of this paper is as follows: @sec-data discusses the data types included in the raw data, the cleaning process for the data, and the reason for selecting the data set we did. @sec-model discusses model specification and justification for a Linear Regression model. @sec-results analyzes the trends and correlations between different variables utilizing tabular and graphical means. @sec-discussion discusses the results of @sec-results going into detail on what the simulation results can tell us about Pakistan cricket performance, as well as assumptions and limitations in data.

## Estimand
The estimand is the probability that Pakistan wins a cricket match, given the opponent, toss outcome, and year.

# Data {#sec-data}

The data for this study were obtained from the cricketdata R package [@cricketdata] and all analysis was performed using `R` [@citeR] alongside the following packages: `tibble` [@tibble], `readr` [@readr], `arrow` [@arrow], `tidyverse` [@tidyverse], `dplyr` [@dplyr], `testthat` [@testthat], `caret` [@caret], `rstanarm` [@rstanarm], `tidyr` [@tidyr], patchwork [@patchwork], and ggplot2 [@ggplot2].

As mentioned before the raw data was collected from the cricketdata library which contains data on international and other major cricket matches sourced from ESPN Cricinfo [@espncricinfo] and Cricsheet [@cricsheet]. It contains details on the individual match year, teams, result, and toss winner. Additionally, the data set features comprehensive career statistics for individual players, including total runs, total balls faced, total wickets, and more. This data set was chosen for its ease of implementation within R, as it is available as a package, and for its ability to scrape data from multiple reputable cricket sources, such as ESPN, thereby enhancing the overall reliability of the data set.

## Raw Data
The cricketdata library offers a variety of functions that provide data sets on different cricket statistics. However, for our analysis, we focus on two key areas: individual match statistics, to determine which matches Pakistan played in and how they performed, and player career statistics, to assess how the quality of players has evolved over time and its impact on Pakistan's overall performance. To extract the relevant data, we primarily used two functions from the library: fetch_cricinfo, which retrieves individual career performance data, and fetch_cricsheet, which provides detailed match information.

## Data Cleaning
The raw match data initially contained 820 observations across 25 variables, each detailing a match between two countries. The data cleaning process involved selecting only matches in which Pakistan played, and retaining only the variables relevant to our analysis. These variables include:

- `Team1` - The first team that participated in the match.
- `Team2` - The second team that participated in the match.
- `Date` - The date of the match (eg. 2008-01-02)
- `Winner` - The team that won the match between Team1 and Team2 (NA if the match was a draw).
- `Winner_wickets` - The number of wickets the winning team took (NA if the match was a draw).
- `Winner_runs` - The number of runs the winning team scored (NA if the match was a draw).
- `Toss_winner` - The team that won the coin toss.
- `Toss_decision` - The decision made by the toss winner, indicating whether they chose to field or bowl.

The raw player career data was broken into batting, bowling, and fielding categories with some players appearing in multiple data sets, thus the cleaning process involved combining the 3 data sets and ensuring proper handling of one player in each set. After this, only Pakistani cricket players were selected and variables which were important to use were retained. These variables are:

- `Player` - Name of the player.
- `Start` - The year in which the player began their career
- `End` - The year in which the player's career concluded.
- `Matches` - The total number of matches the player participated in throughout their career.
- `Bat_innings` -  The total number of batting innings the player participated in.
- `Bowl_innings` -  The total number of bowling innings the player participated in.
- `Field_innings` -  The total number of fielding innings the player participated in.
- `Bat_runs` - The total number of runs the player scored while batting.
- `Not_puts` - The number of times the player remained not out while batting.
- `Bat_average` - The player's batting average, calculated as the total number of runs scored divided by the number of times they were dismissed.
- `Bowl-runs` - The total number of runs conceded by the player while bowling. This statistic reflects how many runs the player allowed the opposition to score during their bowling spells.
- `Wickets` - The total number of wickets taken by the player in their career.
- `Economy` - The average number of runs the player concedes per over bowled. It is calculated by dividing the total number of runs given by the total number of overs bowled. A lower economy rate typically indicates a bowler who is more effective at restricting the opposition's scoring.
- `Dismissals` - The total number of dismissals the player has contributed to as a fielder or wicketkeeper.

After applying these cleaning procedures, we were left with 151 observations across 7 variables for the matches Pakistan has played in, and 344 observations across 14 variables for the career information of individual Pakistani cricketers.

## Measurement

The cricketdata package extracts match information from official sources, including the ESPN and Cricsheet databases, scorecards, and records maintained by cricket boards. These sources provide detailed records of each international match played by Pakistan and other cricketing teams worldwide. These matches represent the real-world phenomena we aim to analyze.

This real-world data is transformed into structured entries in our dataset through the manual recording of match information by officials during the game. Key details, such as match ID, participating teams, venue, date, event, toss winner/decision, player of the match, umpires, match winner, and more, are carefully documented. Once recorded, this information is made available for download on official websites. The datasets are then retrieved and imported into R using the cricketdata package.

This measurement approach is highly reliable, as it relies on official cricketing records. However, it also highlights the limitations inherent in working with aggregated datasets. For example, factors that can influence match outcomes—such as team morale or crowd effects—remain unquantified.

## Variable Analysis
Figures @fig-match-stats-1 and @fig-match-stats-2 display the count of several important match statistics. From these, we can observe that Pakistan faces a variety of opponents, with the most frequent ones being England, Sri Lanka, and Australia. Notably, Pakistan has had a challenging record against Australia, losing a significant number of matches, while many of their matches ended in draws.

Regarding coin toss outcomes, Sri Lanka and South Africa appear to have won a substantial number of tosses. However, it's important to note that while the graphs for match outcomes and tosses won may give the impression that Pakistan was victorious in most of these events, we must keep in mind that the data has been filtered to include only Pakistan's matches. This means Pakistan appears more frequently in the data set compared to other teams, which accounts for the higher statistics shown. These figures represent Pakistan's performance specifically, not that of other teams.

The following two graphs show the number of matches and tosses that Pakistan won or lost. The win-loss ratio is approximately 0.5, indicating that Pakistan has lost about twice as many matches as it has won. Additionally, the coin toss win ratio is close to even, reflecting a near 50/50 chance of Pakistan winning the toss.

Finally, the last graph illustrates the number of matches Pakistan played each year. We can see a peak in 2016, with a gradual decline in matches played over the subsequent years, culminating in a sharp drop in the most recent year—comparable to the period during the COVID-19 pandemic.

```{r}
#| label: fig-match-stats-1
#| fig-cap: Counts for match variables of interst
#| echo: false

data <- pakistan_match_data %>%
  mutate(pakistan_win = case_when(
    winner == "Pakistan" ~ 1,
    TRUE ~ 0))%>%
  mutate(toss_win = case_when(
    toss_winner == "Pakistan" ~ 1,
    TRUE ~ 0))%>%
  mutate(year = lubridate::year(date))

# Convert relevant columns to factors if they represent categorical data
data <- data %>%
  mutate(
    team1 = as.factor(team1),
    team2 = as.factor(team2),
    winner = as.factor(winner),
    toss_winner = as.factor(toss_winner),
    pakistan_win = as.factor(pakistan_win),
    toss_win = as.factor(toss_win)
  )

# Figure 2: Distribution of matches played by Team2
fig2 <- ggplot(data, aes(x = team2)) +
  geom_bar(fill = "coral") +
  labs(title = "Pakistan Opponent Team", x = "Team", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Figure 3: Match outcomes (Winner)
fig3 <- ggplot(data, aes(x = winner)) +
  geom_bar(fill = "darkgreen") +
  labs(title = "Matches Won by Each Team", x = "Team", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Figure 4: Distribution of Toss Winners
fig4 <- ggplot(data, aes(x = toss_winner)) +
  geom_bar(fill = "purple") +
  labs(title = "Tosses Won by each Team", x = "Team", y = "Count") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Figure 5: Pakistan Win Outcomes
fig5 <- ggplot(data, aes(x = pakistan_win)) +
  geom_bar(fill = "gold") +
  labs(title = "Matches Won by Pakistan", x = "Outcome", y = "Count") +
  scale_x_discrete(labels = c("0" = "Lost", "1" = "Won")) +  # Change 1 and 0 to Won and Lost
  theme_minimal()

# Figure 6: Toss Win Outcomes
fig6 <- ggplot(data, aes(x = toss_win)) +
  geom_bar(fill = "dodgerblue") +
  labs(title = "Tosses Won by Pakistan", x = "Outcome", y = "Count") +
  scale_x_discrete(labels = c("0" = "Lost", "1" = "Won")) +  # Change 1 and 0 to Won and Lost
  theme_minimal()

# Figure 7: Match Results by Year
fig7 <- ggplot(data, aes(x = year)) +
  geom_bar(fill = "brown") +
  labs(title = "Pakistan's Matches Played \nper Year", x = "Year", y = "Number of Matches") +
  theme_minimal()

combines_plot1 <- (fig2)/(fig3)
print(combines_plot1)

```

```{r}
#| label: fig-match-stats-2
#| fig-cap: Counts for match variables of interst
#| echo: false

combines_plot2 <- (fig4 | fig5) /( fig6 | fig7)
print(combines_plot2)
```

# Model {#sec-model}

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# library(rstanarm)
# 
# first_model <-
#   readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
# 
# modelsummary::modelsummary(
#   list(
#     "First model" = first_model
#   ),
#   statistic = "mad",
#   fmt = 2
# )
```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

Please don't use these as sub-heading labels - change them to be what your point actually is.

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
# 
# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
```

\newpage

# References