### 2024-12-04 other/model_card/datasheet_template.qmd
Final submission - added discussion, model card, and appendix

---
title: "Model Card for Predicting Pakistan Cricket Match Outcomes"
author: 
  - Muhammad Abdullah Motasim
thanks: "Model can me found at model/win_chance.rds and scripts/05-model_data.R"
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
format: pdf
number-sections: true
---

# Model Card for Predicting Pakistan Cricket Match Outcomes

## Model Details
- **Model Name**: Pakistan Cricket Match Outcome Predictor  
- **Algorithm**: Logistic Regression using Bayesian Generalized Linear Model (via `stan_glm`)  
- **Purpose**: To predict the likelihood of Pakistan winning a cricket match based on historical match data.  
- **Version**: 1.0  
- **Developed By**: Muhammad Abdullah Motasim  
- **Frameworks Used**: R (`rstanarm`, `dplyr`)

## Intended Use
- **Primary Application**:  
  - Predict match outcomes for Pakistan cricket matches.  
  - Analyze factors influencing match results, such as opponent team, toss outcomes, and time trends.  
- **Intended Users**: Cricket analysts, statisticians, fans, and sports strategists.  
- **Use Cases**:  
  - Enhancing cricket strategies by identifying key predictors of success.  
  - Historical performance analysis of Pakistan’s cricket team.

## Model Metrics
- **Training Data Accuracy**: Not explicitly calculated; the model was fit on 80% of the data split.  
- **Testing Data Accuracy**: 82% (rounded from 0.82).  
- **Confusion Matrix**:  
  - **True Positives**: Correctly predicted Pakistan wins.  
  - **True Negatives**: Correctly predicted losses.  
  - **Errors**: Misclassifications in either class.  
- **Model Coefficients**:  
  - **Year**: `0.03 ± 0.04` (indicates a minor increasing trend in win probability over time).  
  - **Opponent Teams**:  
    - Strongest predictors:  
      - **Bangladesh** (`3.58 ± 1.35`) indicates a higher probability of wins.  
      - **West Indies** (`2.20 ± 0.86`) also shows a strong win probability.  
      - **Zimbabwe** (`2.47 ± 1.51`).  
    - Moderate predictors: **England**, **Sri Lanka**.  
    - Weak/uncertain predictors: **India**, **South Africa**, **New Zealand**.  
  - **Toss Win**: `0.59 ± 0.43` (winning the toss slightly increases the probability of winning the match).

## Training Data
- **Source**:  
  - Player and match data files: `cleaned_player_data.parquet`, `cleaned_match_data.parquet`, and `cleaned_pakistan_match_data.parquet`.  
- **Structure**: Includes match-level variables such as winner, toss winner, opponent team, and match year.  
- **Size**: Split into training (80%) and testing (20%) datasets.  
- **Preprocessing**:  
  - Filtering for Pakistan matches.  
  - Encoding binary variables (e.g., match win/loss, toss win/loss).

## Ethical Considerations
- **Bias**: The model's coefficients reveal some bias towards opponent teams (e.g., higher probabilities against weaker teams like Bangladesh and Zimbabwe). This reflects historical trends but might perpetuate biases in cricket analysis.  
- **Data Limitations**:  
  - The dataset does not include contextual factors like weather, player fitness, or ground conditions, which could influence outcomes.  
  - Matches against less frequent opponents may lead to inflated coefficients (e.g., Ireland).  
- **Fairness**: The model does not account for changes in cricket rules or playing conditions over time.  
- **Transparency**: Coefficients and assumptions are clearly documented to ensure interpretability.

## Caveats and Recommendations
- **Caveats**:  
  - The model assumes linear relationships between predictors and outcomes, which may oversimplify complex dynamics in cricket matches.  
  - The data includes only matches where Pakistan participated, limiting generalizability to other teams.  
  - The "toss win" variable shows moderate predictive power, but its causal relationship with match outcomes is not guaranteed.  
- **Recommendations**:  
  - Include additional contextual features (e.g., match location, player statistics) for better predictions.  
  - Validate the model on more recent datasets to ensure reliability over time.  
  - Be cautious when using the model for critical decision-making, as it simplifies cricket's multifaceted nature.

## Acknowledgments
This model was inspired by Pakistan’s rich cricketing history and aims to provide insights into factors affecting their performance. The Bayesian approach ensures probabilistic interpretability and robustness in predictions.

## Contact Information
- **Developer**: Muhammad Abdullah Motasim  
- **Email**: abdullah.motasim@mail.utoronto.ca 
- **GitHub Repository**: https://github.com/abdullah-motasim/Analyzing-Pakistans-Cricket-Data/tree/main